Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2019-10-07 manuscript/manuscript.tex
edits

 Rapid phenotypic adaptation is often observed in natural populations and
 selection experiments.  However, detecting the impact of this selection
 genome-wide is difficult, since adaptation often proceeds from standing
 variation and selection on highly polygenic traits, both of which may leave
 faint genomic signals indistinguishable from a noisy background of genetic
 drift.  One promising signal comes from the genome-wide covariance between
 allele frequency changes observable from temporal genomic data, e.g.
 evolve-and-resequence studies. These temporal covariances reflect how the
 change in neutral allele frequency at one timepoint is predictive of the
 changes at later timepoints when there is heritable fitness variation in the
 population, as neutral alleles can remain associated with selected alleles
 over time. Since genetic drift does not lead to temporal covariance, we can
 use these covariances to estimate what fraction of the variation in allele
 frequency change through time is driven by linked selection. Here, we
 reanalyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
 one artificial selection experiment in mice, to quantify the effects of linked
 selection over short timescales using covariance among time-points and across
 replicates. We estimate that over \vb{TO FIX PUT RANGE XXX} 20\% of allele
 frequency change in these experiments is driven by selection in some
 experiments. Against this background of positive genome-wide temporal
 covariances we also identify signals of negative temporal covariance
 corresponding to reversals in the direction of selection for a reasonable
 proportion of loci over the time course of a selection experiment.  Overall,
 we find that in the three studies we analyzed, linked selection has a large
 impact on short-term allele frequency dynamics that is readily distinguishable
 from genetic drift.

@@

A long-standing problem in evolutionary genetics is quantifying the roles of
genetic drift and selection in shaping genome-wide allele frequency changes.
Selection can both directly and indirectly affect allele frequencies, with the
indirect effect coming from the action of selection on correlated loci
elsewhere in genome e.g. linked selection (\cite{Maynard_Smith1974-lc},
\cite{Charlesworth1993-gb,Nordborg1996-nq}; see \cite{Barton2000-zg} for a
review). Previous work on this question has mostly focused on teasing apart
the impacts of drift and selection on genome-wide diversity using population
samples from a single contemporary timepoint, often by modeling the correlation
between regional recombination rate, gene density, and diversity created in the
presence of linked selection \parencite{SellaReview2009,CutterPayseur}. This  %\parencite{Begun1992-ey,Elyashiv2016-vt}
approach has shown linked selection has a major role in shaping patterns of
genome-wide diversity across the genomes of \gc{a range of sexual species}
\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-v},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and negative selection (background selection;
\cite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt}.
However, we lack an understanding of how genome-wide linked selection acts over
time.

There are numerous examples of rapid phenotypic adaptation
\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid
\gc{, selection-driven} genomic evolution in asexual populations
\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
of fitness makes detecting the impact selection on genome-wide variation over
short timescales in sexual populations remarkably difficult. This is because
the effect of selection on a polygenic trait (such as fitness) is distributed
across loci in proportion to their effect sizes. This can lead to subtle allele
frequency shifts on standing variation that are difficult to distinguish from
background levels of genetic drift and sampling variance.  However,
increasingly genomic experimental evolution studies with multiple timepoints,
and in some cases multiple replicate populations, are being used to detect
\gc{large effect} selected loci \parencite{Turner2011-sx,Turner2012-bm} and differentiate modes
of selection \parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In
addition these temporal-genomic studies have begun in wild populations, some
with the goal of finding variants that exhibit frequency changes consistent
with fluctuating selection \parencite{Bergland2014-ij,Machado2018-cs}. In a
previous paper, we proposed that one useful signal for understanding the
genome-wide impact of polygenic linked selection detectable from temporal
genomic studies is the temporal autocovariance in allele frequency changes
\parencite{Buffalo2019-io}.  These covariances are directly estimable from
temporal genomic data and are created when the loci that underly heritable
fitness variation perturb the frequencies of linked neutral alleles; in
contrast, when genetic drift acts alone in a closed population, these
covariances are zero in expectation. Mathematically, temporal covariances are
useful because it is natural to decompose the total variance in allele
frequency change across a set of time intervals into the variances and
covariances in allele frequency change among time intervals. Furthermore,
biologically, these covariances reflect the extent to which allele frequency
changes in one generation predict changes in another due to a shared \gc{genetic} %fitness
background and \gc{correlated} selection pressures. %shared

Here, we provide the first empirical analyses to quantify the impact of linked
selection acting over short timescales (tens of generations) across two evolve
and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and \gc{an}
artificial selection experiment \parencite{Castro2019-uk}. We repeatedly find a
signal of temporal covariance, consistent with linked selection acting to
significantly perturb genome-wide allele frequency changes across the genome in
a manner that other approaches would not be able differentiate from genetic
drift. We estimate the lower bound on the proportion of total variation in
allele frequency change caused by \gc{selection}, and the correlation
between  %linked
allele frequency changes between replicate populations caused by response to
convergent selection pressures. Overall, we demonstrate that linked selection
has a powerful role in shaping genome-wide allele frequency changes over very
short timescales.

@@

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature lab environment and
evolved for 60 generations, and sequenced every ten generations. Using the
seven timepoints and ten replicate populations, we estimated the genome-wide $6
\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
replicates. Each row of these matrices represent the temporal covariance
$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
change (in ten-generation intervals, denoted $\Delta_{_{10}} p_t$) in some
initial reference generation $s$ (the row of the matrix), and some later
timepoint $t$ (the column of the matrix). We corrected these matrices for
biases created due to sampling noise, and normalize the entries for
heterozygosity (see Supplementary Material Sections
\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction} for details on
the bias correction). The covariances are expected to be zero when only drift
is acting, as only heritable variation for fitness can
create covariance between allele frequency changes \gc{in a closed population} \parencite{Buffalo2019-io}.
Averaging across the ten replicate temporal covariances matrices, we find
temporal covariances that are statistically significant (95\% block bootstraps
CIs do not contain zero), consistent with linked selection perturbing
genome-wide allele frequency changes over very short time periods. The
covariances between all adjacent time intervals are positive and then decay
towards zero as we look at more distant time intervals, as expected when
directional selection affects linked variants' frequency trajectories until
ultimately linkage disequilibrium and the additive genetic variance for fitness
associated with neutral alleles decays \parencite{Buffalo2019-io}. The temporal
covariances per replicate are noisier but this general pattern holds; see
Supplementary Figures \ref{suppfig:barghi-cov-panels}.
\textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
time intervals are on average ten generations apart, and given the temporal
decay in covariance we see, the covariances on shorter time-scales (e.g. if
adjacent generations had been sequenced) may well be higher yet (see
Supplementary Material Section \ref{supp:barghi-covs} for more details).

\gc{One concern is that these covariances reflect the localized impact
  of a few large-effect loci
  rather than selection on a polygenic trait. Since our covariances are
  essentially averages over loci, the covariance estimate can be strongly
  affected by a single outlier region. To test whether large outlier regions
  drive the genome-wide signal we see in the \textcite{Barghi2019-qy} data, we
  calculate the covariances in 100kb windows along the genome (we refer to
  these as windowed covariances throughout). We take the median
  windowed covariance as a measure of the genome-wide covariance robust to
  large-effect loci; we find that the median windowed covariance $\cov(\Delta
  p_0, \Delta p_{10})$ (pooling all replicates) is also significantly different
  from zero, 0.0016 (0.95\% CIS [0.0015, 0.0017]). This median
  covariance is also limiting result of a trimmed mean that excludes the upper and lower
  $\alpha$ tails to calculate the trimmed average windowed covariance. As
  $\alpha$ increases to 0.5, the trimmed covariance converges to the median
  windowed covariance (by the definition of the median; see Supplementary
  Figure \ref{suppfig:barghi-trimmed-mean}). Thus our genomic temporal
  covariances are non-zero due to the impact of selection on many
  genomic windows.
  XXX put a table median for all timepoints. AND REWORK TEXT
 }

   \end{tabular}

   \caption{Table of median of windowed covariance estimates ($\cov(\Delta p_s,
     \Delta p_t) \times 100$) between generations $t$ and $s$ and the trimmed
     mean windowed covariance which excludes the lower and upper 5\% windows
     with the highest covariance.}

@@

While the presence of positive temporal covariances is consistent with 
\gc{selection} affecting allele frequencies over time, this measure not easily
interpretable. \gc{We can calculate a more intuitive measure from this
number to} quantify the impact of \gc{selection} on
allele frequency change as the ratio of total covariance in allele frequency
change to the total variance in allele frequency change. We denote the
change in allele frequency as $\Delta p_t = p_{t+1}-p_t$, where $p_t$ is the
allele frequency in generation $t$. Since the total variation in allele
frequency change can be partitioned into variance and covariance components,
$\var(p_t - p_0) = \sum_{i} \var(\Delta p_i) + \sum_{i \ne j} \cov(\Delta
p_i, \Delta p_j)$, and the covariances are zero when drift acts alone, this is
a lower bound on how much of the variance in allele frequency change is caused
by linked selection \parencite{Buffalo2019-io}. We call this measure $G(t)$, defined as

\begin{align}
  G(t) = \frac{\sum_{i\ne j}^t \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}
\end{align}
%
which estimates the effect of \gc{selection on allele frequency
  change} between the initial
generation $0$ and some later generation $t$, which can be varied to see how
this quantity grows through time. As with the temporal covariances, the study
design of \textcite{Barghi2019-qy} leads our measure $G(t)$ to be even more
conservative, since the temporal covariances within each ten-generation block
between sequenced timepoints are not directly observable, and are not included
in the numerator of $G(t)$. Still, we find a remarkably strong signal.
Greater than $20\%$ of total\gc{, genome-wide allele frequency
  change} over 60 %variation in
generations is the result of selection.

@@

\vb{We note that blah blha lbah, have a look at the pairwiwe covariances for
Fig 2A} \gc{THIS IS about double checking that that last time point is
about right?}

@@

We've calculated the convergent correlation for all rows of the replicate
covariance matrices. Like temporal covariances, we visualize these through time
(Figure \ref{fig:figure-2} A), with each line representing the convergent
correlation from a particular reference generation $s$ as it varies with $t$
(shown on the x-axis). In other words, each of the colored lines corresponds to
the like-colored row of the convergence correlation matrix (upper left in
Figure \ref{fig:figure-2} A). We find these convergent covariances are
relatively weak, and decay very quickly from an initial value of about 0.1
(95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
(95\% CIs [0.0087, 0.015]) within 20 generations. This \gc{is
  consistent with the}
primary finding of the original \textcite{Barghi2019-qy} study: that
alternative loci contribute to \gc{longer term} adaptation in different replicates.

@@

A benefit of between-replicate covariances is that unlike temporal covariances,
these can be calculated with only two sequenced timepoints and a replicated
study design. This allowed us to assess the impact of linked selection in
driving convergent patterns of allele frequency change across replicate
populations in two other studies. First, we reanalyzed the selection experiment
of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
environment. Since each replicate was exposed to the same selection pressure
and share linkage disequilibria common to the original natural founding
population, we expected each of the three replicate populations to have
positive between-replicate covariances. We find all three pairwise
between-replicate covariances are positive and statistically significant (95\%
CIs, [0.0063, 0.0086], [0.0064, 0.0086], [0.0061, 0.0083]). We estimate the
convergent correlation coefficient across these replicates as 0.36 ($95\%$
block-bootstrap confidence interval $[0.31, 0.40]$). \vb{Put in a note about
replicate G.}

@@

\vb{The presence of an unselected control line provides an alternative way to
  partition the effects of linked selection and genetic drift: we can compare
  the total

}

@@

% Here, we extend our theoretic work in \textcite{Buffalo2019-io} to
% between-replicate allele frequency change covariances. We focus on the
% between-replicate covariance at a focal neutral site across two replicate
% populations, $A$ and $B$. As in \textcite{Buffalo2019-io}, the heritable allele
% frequency change (which is the only term that contributes to the covariance),
% can be written as a linear combination of effect sizes $\alpha_{t,l,A}$ and LD
% between the neutral site and selected site $l$, $D_{t,l,A}$.

% \begin{align}
%   \cov(\Delta p_{t,A}, \Delta p_{t,B}) = \cov\left(\sum_l \alpha_{t,A,l} D_{t,A,l}, \sum_l \alpha_{t,B,l} D_{t,B,l}\right)
% \end{align}

% Since the replicate populations derive from a common population at time $t=0$
% before their split, they share a fraction of their LD plus a replicate-specific
% deviation due to replicate-specific selection and drift. Thus, we can partition
% the $D$ terms as: $D_{t,A,l} = D_{0,l} + \varepsilon_{t,A}$ (and likewise for
% replicate $B$). Under drift alone, this deviation can be found from equation
% (4) of \textcite{Ragsdale2019-fs}.  Similarly, in the original population the
% additive effects at each locus were identical ($\alpha_{0,l}$), but in each
% replicate population have since been perturbed in a replicate-specific manner
% (the term $a_{t,A,l}$) as well as a change induced by convergent selection
% pressure after $t$ generations, $b_{t,l}$, $\alpha_{0,l} + b_{t,l} + a_{t,A,l}$
% (and likewise for replicate $B$).

% We then can write,

% \begin{align}
%   \cov(\Delta p_{t,A}, \Delta p_{t,B}) = \cov\left(\sum_l \alpha_{t,A,l} (D_{0,l} + \varepsilon_{A,l}), \sum_l \alpha_{t,B,l} (D_{0,l} + \varepsilon_{B,l})\right)
% \end{align}

% We define a statistic similar to $G$ for estimating the proportion of allele
% frequency change common between two replicate populations due to linked
% selection. Covariance in allele frequency change between two replicate
% populations is due to convergent selection pressure selecting haplotypes shared
% between the two replicate populations, which acts to perturb linked neutral
% variation in parallel way. Similar to \textcite{Buffalo2019-io} equation (1).
% We can similarly partition the allele frequency change,

% \begin{align}
%   \label{eq:shared-partition}
%   \Delta p_t &= \underbrace{\Delta_{_N} p_t  + \Delta_{_M} p_t}_\text{drift} + \underbrace{\Delta_{_S} p_t}_\text{shared} + \underbrace{\Delta_{_U} p_t}_\text{unique}
% \end{align}

% where $\Delta_{_N} p_t$ and $\Delta_{_M} p_t$ are the allele frequency changes
% due to non-heritable variation in fitness (e.g. environmental differences in
% family size) and Mendelian segregation. In \textcite{Buffalo2019-io} allele
% frequency changes due to heritable fitness variation are captured in the term
% $\Delta_{_H} p_t$; here we further partition this into heritable frequency
% changes shared between replicates ($\Delta_{_S} p_t$) and those unique to a
% replicate ($\Delta_{_H} p_t$). This is important, as we can

% We base this partition on our model in
% \textcite{Buffalo2019-io}, where we decompose the LD between the 
% focal neutral site and selected sites conditioning on whether

% \begin{align}
%   \Delta_{_H} p_{t,A} &= \sum_l \alpha_{t,l,A} D_{t,l, A} \\
%                       &= \alpha_{t,l,A} D_{t,l,A} + (\alpha_{t,l,A'} + \alpha_{t,l,S}) D_{t,l,S} \\
% \end{align}

% By definition,

% \begin{align}
%   \cov(\Delta p_{s,A}, \Delta p_{t,B}) &= \cov(\Delta_{_S} p_{s,A}, \Delta_{_S} p_{t,B}), \;\; A \ne B\\
%   \cov(\Delta p_{s,A}, \Delta p_{t,A}) &= \cov(\Delta_{_S} p_{s,A} + \Delta_{_U} p_{s,A}, \Delta_{_S} p_{t,A} + \Delta_{_U} p_{t,A}) 
% \end{align}

% For selection studies with a control line (e.g. the \cite{Castro2019-uk} study
% we have reanalyzed), we have an alternate approach to quantify the effect of
% linked selection: we can compare the total variance in allele frequency change
% in control and selection lines. By the orthogonality of the components in
% equation XXX,

% \begin{align}
%   \var(\Delta p_t - p_0) = \sum_t \var(\Delta {_N} p_t ) + \sum_t \var(\Delta {_M} p_t ) + 
%   \var\left(\sum_t \Delta {_S} p_t\right) + \var\left(\sum_t \Delta {_U} p_t\right)
% \end{align}

% the term $\sum_t \var(\Delta {_S} p_t)$

% \begin{align}
%   \var\left(\sum_t \Delta_{_S} p_t\right) = 
% \sum_t \var( \Delta_{_U} p_t) + \sum_{i \ne j} \cov( \Delta_{_S} p_i, \Delta_{_S} p_j)
% \end{align}

% In the \textcite{Castro2019-uk} study, we 
% quantify the effect of linked selection as:

% \begin{align}
%   \frac{1}{2} \left( \var(\Delta p_{20, \mathrm{LS1}} - p_{0, \mathrm{LS1}}) + \var(\Delta p_{20, \mathrm{LS2}} - p_{0, \mathrm{LS2}}) \right) -  \var(\Delta p_{20, \mathrm{Ctrl}} - p_{0, \mathrm{Ctrl}}) = \\
%   \cov(p_{20, \mathrm{LS1}} - p_{0, \mathrm{LS1}}, p_{20, \mathrm{LS2}} - p_{0, \mathrm{LS2}})
% \end{align}

% \begin{align}
%   \var(p_t - p_0) &= \sum_i \var(\Delta p_i) + \sum_{i\ne j} \cov(\Delta p_i, \Delta p_j) \\
%                   &= \sum_i \var(\Delta p_i) + \sum_{i\ne j} \cov(\Delta_{_U} p_i, \Delta_{_U} p_j) + \sum_{i\ne j} \cov(\Delta_{_S} p_i, \Delta_{_S} p_j)  \\
% \end{align}

% However, with only two timepoints, we cannot directly measure the unique
% temporal covariance. However, we can directly measure the covariance between
% replicates $A$ and $B$, $\sum_{i,j} \cov(\Delta p_{i,A}, \Delta p_{j,B})$.

@@

>>>>

 Rapid phenotypic adaptation is often observed in natural populations and
 selection experiments.  However, detecting the impact of this selection
 genome-wide is difficult, since adaptation often proceeds from standing
 variation and selection on highly polygenic traits, both of which may leave
 faint genomic signals indistinguishable from a noisy background of genetic
 drift.  One promising signal comes from the genome-wide covariance between
 allele frequency changes observable from temporal genomic data, e.g.
 evolve-and-resequence studies. These temporal covariances reflect how the
 change in neutral allele frequency at one timepoint is predictive of the
 changes at later timepoints when there is heritable fitness variation in the
 population, as neutral alleles can remain associated with selected alleles
 over time. Since genetic drift does not lead to temporal covariance, we can
 use these covariances to estimate what fraction of the variation in allele
 frequency change through time is driven by linked selection. Here, we
 reanalyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
 one artificial selection experiment in mice, to quantify the effects of linked
 selection over short timescales using covariance among time-points and across
 replicates. We estimate that over \vb{17\% to 37\%} of allele
 frequency change in these experiments is driven by selection in some
 experiments. Against this background of positive genome-wide temporal
 covariances we also identify signals of negative temporal covariance
 corresponding to reversals in the direction of selection for a reasonable
 proportion of loci over the time course of a selection experiment.  Overall,
 we find that in the three studies we analyzed, linked selection has a large
 impact on short-term allele frequency dynamics that is readily distinguishable
 from genetic drift.

@@

A long-standing problem in evolutionary genetics is quantifying the roles of
genetic drift and selection in shaping genome-wide allele frequency changes.
Selection can both directly and indirectly affect allele frequencies, with the
indirect effect coming from the action of selection on correlated loci
elsewhere in genome e.g. linked selection (\cite{Maynard_Smith1974-lc},
\cite{Charlesworth1993-gb,Nordborg1996-nq}; see \cite{Barton2000-zg} for a
review). Previous work on this question has mostly focused on teasing apart the
impacts of drift and selection on genome-wide diversity using population
samples from a single contemporary timepoint, often by modeling the correlation
between regional recombination rate, gene density, and diversity created in the
presence of linked selection \parencite{Cutter2013-ba,Sella2009-nx}. This
approach has shown linked selection has a major role in shaping patterns of
genome-wide diversity across the genomes of a range of sexual species
\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and negative selection (background selection;
\cite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt}.
However, we lack an understanding of how genome-wide linked selection acts over
time.

There are numerous examples of rapid phenotypic adaptation
\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid,
selection-driven genomic evolution in asexual populations
\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
of fitness makes detecting the impact selection on genome-wide variation over
short timescales in sexual populations remarkably difficult. This is because
the effect of selection on a polygenic trait (such as fitness) is distributed
across loci in proportion to their effect sizes. This can lead to subtle allele
frequency shifts on standing variation that are difficult to distinguish from
background levels of genetic drift and sampling variance.  However,
increasingly genomic experimental evolution studies with multiple timepoints,
and in some cases multiple replicate populations, are being used to detect
large effect selected loci \parencite{Turner2011-sx,Turner2012-bm} and
differentiate modes of selection
\parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In addition these
temporal-genomic studies have begun in wild populations, some with the goal of
finding variants that exhibit frequency changes consistent with fluctuating
selection \parencite{Bergland2014-ij,Machado2018-cs}. In a previous paper, we
proposed that one useful signal for understanding the genome-wide impact of
polygenic linked selection detectable from temporal genomic studies is the
temporal autocovariance in allele frequency changes \parencite{Buffalo2019-io}.
These covariances are directly estimable from temporal genomic data and are
created when the loci that underly heritable fitness variation perturb the
frequencies of linked neutral alleles; in contrast, when genetic drift acts
alone in a closed population, these covariances are zero in expectation.
Mathematically, temporal covariances are useful because it is natural to
decompose the total variance in allele frequency change across a set of time
intervals into the variances and covariances in allele frequency change among
time intervals. Furthermore, biologically, these covariances reflect the extent
to which allele frequency changes in one generation predict changes in another
due to a shared genetic pressures.

Here, we provide the first empirical analyses to quantify the impact of linked
selection acting over short timescales (tens of generations) across two evolve
and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and an
artificial selection experiment \parencite{Castro2019-uk}. We repeatedly find a
signal of temporal covariance, consistent with linked selection acting to
significantly perturb genome-wide allele frequency changes across the genome in
a manner that other approaches would not be able differentiate from genetic
drift. We estimate the lower bound on the proportion of total variation in
allele frequency change caused by selection, and the correlation between
allele frequency changes between replicate populations caused by response to
convergent selection pressures. Overall, we demonstrate that linked selection
has a powerful role in shaping genome-wide allele frequency changes over very
short timescales.

@@

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature lab environment and
evolved for 60 generations, and sequenced every ten generations. Using the
seven timepoints and ten replicate populations, we estimated the genome-wide $6
\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
replicates. Each row of these matrices represent the temporal covariance
$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
change (in ten-generation intervals, denoted $\Delta_{_{10}} p_t$) in some
initial reference generation $s$ (the row of the matrix), and some later
timepoint $t$ (the column of the matrix). We corrected these matrices for
biases created due to sampling noise, and normalize the entries for
heterozygosity (see Supplementary Material Sections
\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction} for details on
the bias correction). The covariances are expected to be zero when only drift
is acting, as only heritable variation for fitness can create covariance
between allele frequency changes in a closed population
\parencite{Buffalo2019-io}.  Averaging across the ten replicate temporal
covariances matrices, we find temporal covariances that are statistically
significant (95\% block bootstraps CIs do not contain zero), consistent with
linked selection perturbing genome-wide allele frequency changes over very
short time periods. The covariances between all adjacent time intervals are
positive and then decay towards zero as we look at more distant time intervals,
as expected when directional selection affects linked variants' frequency
trajectories until ultimately linkage disequilibrium and the additive genetic
variance for fitness associated with neutral alleles decays
\parencite{Buffalo2019-io}. The temporal covariances per replicate are noisier
but this general pattern holds; see Supplementary Figures
\ref{suppfig:barghi-cov-panels}.  \textcite{Barghi2019-qy}'s design means that
the covariances we see in adjacent time intervals are on average ten
generations apart, and given the temporal decay in covariance we see, the
covariances on shorter time-scales (e.g. if adjacent generations had been
sequenced) may well be higher yet (see Supplementary Material Section
\ref{supp:barghi-covs} for more details).

\vb{One concern is that these covariances reflect the localized impact of a few
  large-effect loci rather than selection on a polygenic trait. Since our
  covariances are essentially averages over loci, the covariance estimate can
  be strongly affected by a single outlier region. To test whether large
  outlier regions drive the genome-wide signal we see in the
  \textcite{Barghi2019-qy} data, we calculate the covariances in 100kb windows
  along the genome (we refer to these as windowed covariances throughout). We
  take the median windowed covariance as a measure of the genome-wide
  covariance robust to large-effect loci; we find that the median windowed
  covariance $\cov(\Delta p_0, \Delta p_{10})$ (pooling all replicates) is also
  significantly different from zero, 0.0016 (0.95\% CIS [0.0015, 0.0017], see
  Supplementary Table \ref{supp:table-trimmed-mean} for more timepoints and
  trimmed means). This median covariance is also limiting result of a trimmed
  mean that excludes the upper and lower $\alpha$ tails to calculate the
  trimmed average windowed covariance. As $\alpha$ increases to 0.5, the
  trimmed covariance converges to the median windowed covariance (by the
  definition of the median; see Supplementary Figure
  \ref{suppfig:barghi-trimmed-mean}). Thus our genomic temporal covariances are
  non-zero due to the impact of selection on many genomic windows.  }

@@

While the presence of positive temporal covariances is consistent with
selection affecting allele frequencies over time, this measure not easily
interpretable. \vb{We can calculate a more intuitive measure from the temporal
covariances to quantify the impact of selection on allele frequency change:
the ratio of total covariance in allele frequency change to the total variance
in allele frequency change}. We denote the change in allele frequency as
$\Delta p_t = p_{t+1}-p_t$, where $p_t$ is the allele frequency in generation
$t$. Since the total variation in allele frequency change can be partitioned
into variance and covariance components, $\var(p_t - p_0) = \sum_{i}
\var(\Delta p_i) + \sum_{i \ne j} \cov(\Delta p_i, \Delta p_j)$, and the
covariances are zero when drift acts alone, this is a lower bound on how much
of the variance in allele frequency change is caused by linked selection
\parencite{Buffalo2019-io}. We call this measure $G(t)$, defined as

\begin{align}
  G(t) = \frac{\sum_{i\ne j}^t \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}
\end{align}
%
which estimates the effect of selection on allele frequency change between the
initial generation $0$ and some later generation $t$, which can be varied to
see how this quantity grows through time. As with the temporal covariances, the
study design of \textcite{Barghi2019-qy} leads our measure $G(t)$ to be even
more conservative, since the temporal covariances within each ten-generation
block between sequenced timepoints are not directly observable, and are not
included in the numerator of $G(t)$. Still, we find a remarkably strong signal.
Greater than $20\%$ of total, genome-wide allele frequency change over 60
generations is the result of selection.

@@

@@

We've calculated the convergent correlation for all rows of the replicate
covariance matrices. Like temporal covariances, we visualize these through time
(Figure \ref{fig:figure-2} A), with each line representing the convergent
correlation from a particular reference generation $s$ as it varies with $t$
(shown on the x-axis). In other words, each of the colored lines corresponds to
the like-colored row of the convergence correlation matrix (upper left in
Figure \ref{fig:figure-2} A). We find these convergent covariances are
relatively weak, and decay very quickly from an initial value of about 0.1
(95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
(95\% CIs [0.0087, 0.015]) within 20 generations. This is consistent with
the primary finding of the original \textcite{Barghi2019-qy} study: that
alternative loci contribute to longer term adaptation in different
replicates.  \vb{We note that convergent correlation for the last timepoint
is abnormally high; this is not in fact driven by a single outlier pair of
covariances but all our high for reasons unknown to us.}

@@

A benefit of between-replicate covariances is that unlike temporal covariances,
these can be calculated with only two sequenced timepoints and a replicated
study design. This allowed us to assess the impact of linked selection in
driving convergent patterns of allele frequency change across replicate
populations in two other studies. First, we reanalyzed the selection experiment
of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
environment. Since each replicate was exposed to the same selection pressure
and share linkage disequilibria common to the original natural founding
population, we expected each of the three replicate populations to have
positive between-replicate covariances. We find all three pairwise
between-replicate covariances are positive and statistically significant (95\%
CIs, [0.0063, 0.0086], [0.0064, 0.0086], [0.0061, 0.0083]). We estimate the
convergent correlation coefficient across these replicates as 0.36 ($95\%$
block-bootstrap confidence interval $[0.31, 0.40]$). \vb{Similarly, we can
  calculate the proportion of the total variance in allele frequency change
  from convergent selection pressure analogous to $G$ where the numerator is is
  the convergent covariance and the denominator is the total variance (see
  XXX). We find that 37\% of the total variance is due to shared allele
frequency changes caused by selection (95\% CI [29\%, 41\%]; these are similar
to the convergence correlation, see XXX).}

@@

\vb{The presence of an unselected control line provides an alternative way to
  partition the effects of linked selection and genetic drift: we can compare
  the total variance in allele frequency change of the control line (which
  excludes the effect of artificial selection on allele frequencies) to the
  total variance in frequency change of the Longshanks selection lines. We can
  partition the allele frequency change between the two timepoints (20
  generations) for a Longshanks line as $\Delta p_{t,\mathrm{LS1}} =
  \Delta_{_D} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} +
  \Delta_{_\mathrm{LS}} p_t$ (and similarly for the Longshanks two line, LS2),
  and $\Delta p_{t,\mathrm{Ctrl}} = \Delta_{_D} p_{t,\mathrm{Ctrl}} +
  \Delta_{_{U}} p_{t,\mathrm{Ctrl}} + \Delta_{_\mathrm{Ctrl}} p_t$ for the
  control line. Here, $\Delta_{_D} p_{t}$, $\Delta_{_U} p_t$,
  $\Delta_{_\mathrm{LS}} p_t$ are the allele frequency changes due to drift,
  selection unique to a replicate, and selection response shared between the
  two Longshanks replicates, respectively. By construction each of these terms
  are uncorrelated, allowing us to estimate the increase in variance in allele
  frequency change due to selection as $(\var(\Delta p_{t,\mathrm{LS1}}) +
  \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}})$ and
  the shared effect of selection across selected lines as $\cov(\Delta
  p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$. We estimate at least 32\%
    (95\% CI $[21\%, 48\%]$) of the variance in allele frequency change is
    driven by the effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$)
    is estimated to be unique to a selection line, and 17\% (95\% CI $[9\%,
  23\%]$) is the effect of shared selection between the two Longshanks
selection lines.}

@@

@@

   \end{tabular}
  \label{supp:table-trimmed-mean}
   \caption{Table of median of windowed covariance estimates ($\cov(\Delta p_s,
     \Delta p_t) \times 100$) between generations $t$ and $s$ and the trimmed
     mean windowed covariance which excludes the lower and upper 5\% windows
     with the highest covariance.}
